---
title: "Negative Projection Follow-up 1"
author: "Kathryn Denning"
date: "Originally written June 2018; Updated July 2019"
output: 
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
---
**Variable Glossary:**

*Predictor Variables:* 

Condition: What issue participants were given; 1 = political candidate, 2 = racial discrimination, 3 = weed, 4 = PB

Issue_agreement: 1 for condition 1 is Clinton, 2 is Trump, for other three conditions 1 is yes to agreeing with issue, 4 is no to agreeing with issue (AKA ignoring PB, 1 = liberal, 4 = conservative beliefs)

Gender: 1 = male, 2 = female, and 3 = other

Race: 1 = American Indian, 2 = Asian, 3 = African American, 4 = Native Hawaiian/pacific islander, 5 = Caucasian, 6 = Other

Identification: how highly they identify with others who feel the same about them on the issue; 1 = not at all identified, 5 = extremely identified; Identification_c is mean centered

ManCheck: 5 = correct response (all others already removed)

MFQSelf: Scores on MFQ for themselves/the participants; MFQSelf_c is mean centered

*Outcome Variables:*

MFQTarg: Scores on MFQ for target individual who is in out-group/believes the opposite of them on that issue

```{r set-up, include = FALSE, echo = FALSE}
#Removing code from final output
knitr::opts_chunk$set(echo = TRUE) 

#Loading packages
library(knitr)
library(tidyr)
library(haven) 
library(psych)
library(psy)
library(irr)
library(utils)
library(lme4)
library(ggplot2)
library(car)
library(sjPlot)
library(dplyr)
library(psych)
library(effects)
library(lsmeans)
library(lmerTest)

#Reading in the data
pptnpf1 <- read.csv("PPTNPF1_clean_nopt.csv", header = TRUE) 

#Getting the data in long-format
pptnpf1_comb <-  pptnpf1 %>% 
  dplyr::select(Subject, Condition, Gender, Race, Age, Issue_agreement, 
                Identification, ManCheck, 
                MFQSelf.=MFQ_Self_1:MFQ_Self_23, 
                MFQTarg.=MFQ_Targ_1:MFQ_Targ_23) %>% 
  gather(MFQType, MFQResponse, MFQSelf.1:MFQTarg.22, factor_key = TRUE) %>% 
  separate(.,MFQType,c("MFQ","number"),"\\.") %>% 
  spread(.,MFQ, MFQResponse) %>% 
  mutate_each(.,funs(as.numeric(as.character(.)))) %>% 
  group_by(Subject)

#Making each value the right type (numeric or factor) and centered
pptnpf1_comb$MFQSelf_c <- as.numeric(pptnpf1_comb$MFQSelf - 
                                       mean(pptnpf1_comb$MFQSelf, na.rm = TRUE))
pptnpf1_comb$Identification_c <- as.numeric(pptnpf1_comb$Identification - 
                                              mean(pptnpf1_comb$Identification, 
                                                   na.rm = TRUE))
pptnpf1_comb$Condition <- as.factor(pptnpf1_comb$Condition)
pptnpf1_comb$Gender <- as.factor(pptnpf1_comb$Gender)
pptnpf1_comb$Race <- as.factor(pptnpf1_comb$Race)
pptnpf1_comb$Age <- as.numeric(pptnpf1_comb$Age)
pptnpf1_comb <- na.omit(pptnpf1_comb)

#Adding contrast codes for our categorical variable
PRWvPB <- c(1, -1, -1, 3) 
PRvW <- c(-1, -1, 2, 0)  
PvR <- c(-1, 1, 0, 0) 
ConCodes <- cbind(PRWvPB, PRvW, PvR)
contrasts(pptnpf1_comb$Condition) <- ConCodes
contrasts(pptnpf1_comb$Condition)
```

# Reliability of scales

## Reliability for target MFQ 

```{r reliability targ}
MFQtarg_data <- subset(pptnpf1, select = c(6, 7, 8, 9, 10, 12, 13, 14, 
                                           15, 16, 17, 18, 19, 20, 21, 
                                           23, 24, 25, 26, 27))
cronbach(MFQtarg_data)
```

## Reliability for MFQ for participant (self)

```{r reliability self}
MFQself_data <- subset(pptnpf1,select = c(MFQ_Targ_1, MFQ_Targ_2, MFQ_Targ_3, 
                                        MFQ_Targ_4, MFQ_Targ_5, MFQ_Targ_7, 
                                        MFQ_Targ_8, MFQ_Targ_9, MFQ_Targ_10, 
                                        MFQ_Targ_11, MFQ_Targ_12, MFQ_Targ_13, 
                                        MFQ_Targ_15, MFQ_Targ_16, MFQ_Targ_17, 
                                        MFQ_Targ_19, MFQ_Targ_20, MFQ_Targ_21, 
                                        MFQ_Targ_22, MFQ_Targ_23))
MFQself_data <- na.omit(MFQself_data)
cronbach(MFQself_data)
```

# Descriptives

## Descriptives for the variables overall

```{r descriptives overall}
describe(pptnpf1_comb$Identification)
describe(pptnpf1_comb$MFQSelf)
describe(pptnpf1_comb$MFQTarg)
```

## Mean and SD for MFQ

### MFQ for Self

```{r MFQ self}
pptnpf1_comb %>% 
  select(Subject, Condition, MFQSelf) %>% 
  na.omit() %>% 
  group_by(Condition) %>% 
  summarize(mean = mean(MFQSelf),
            sd = sd(MFQSelf))
```

### MFQ for Target

```{r MFQ targ}
pptnpf1_comb %>% 
  select(Subject, Condition, MFQTarg) %>% 
  na.omit() %>% 
  group_by(Condition) %>% 
  summarize(mean = mean(MFQTarg),
            sd = sd(MFQTarg))
```

## Mean and SD for in-group identification

```{r identification}
pptnpf1_comb %>% 
  select(Subject, Condition, Identification) %>% 
  na.omit() %>% 
  group_by(Condition) %>% 
  summarize(mean = mean(Identification),
            sd = sd(Identification))
```

# Demographics

## Number of participants - Overall

```{r n overall}
pptnpf1_comb  %>% 
  select(Subject) %>% 
  unique() %>% 
  nrow()
```

## Number of participants - By condition

```{r n by cond}
pptnpf1_comb %>% 
  select(Subject, Condition) %>% 
  mutate(Condition = recode(Condition,
                            `1` = "Political Candidate",
                            `2` = "Racial Disadvantage",
                            `3` = "Marijuana Legalization",
                            `4` = "Peanut Butter Preference")) %>% 
  group_by(Condition) %>% 
  unique() %>% 
  count()
```

## Number of participants - By gender

```{r n by gender}
pptnpf1_comb %>% 
  select(Subject, Gender) %>% 
    mutate(Gender = recode(Gender,
                         `1` = "Male",
                         `2` = "Female")) %>% 
  group_by(Gender) %>% 
  unique() %>% 
  count()
```

## Number of participants - By race

```{r n by race}
pptnpf1_comb %>% 
  select(Subject, MFQTarg, MFQSelf_c, Identification_c, Condition, Race) %>% 
  na.omit() %>% 
  select(Subject, Race) %>% 
  unique() %>% 
  group_by(Race) %>% 
  count()
```

### Caucasian Percentage

```{r caucasian}
white <- (194/306)*100
white
```

### Asian Percentage

```{r asian}
asian <- (59/306)*100
asian
```

### Identified as "Other" Percentage

```{r other}
other <- (32/306)*100
other
```

### Black Percentage

```{r black}
black <- (14/306)*100
black
```

### Hawaiian/Pacific Islander Percentage

```{r hawaii}
hawaiipacific <- (8/306)*100
hawaiipacific
```

## Age

```{r}
pptnpf1_comb %>% 
  select(Subject, MFQTarg, MFQSelf_c, Identification_c, Condition, Age) %>% 
  na.omit() %>% 
  select(Subject, Age) %>% 
  unique() %>% 
  summarize(mean(Age)) %>% 
  summarize(Mean = mean(`mean(Age)`),
            SD = sd(`mean(Age)`))
```

# Model


## Contrast codes

```{r contrasts}
#contrasts assigned to condition
contrasts(pptnpf1_comb$Condition)
```

* PRWvPB = Politics, Racial discrimination, and Weed legalization vs Peanut butter
* PRvW = Politics and racial discrimination vs weed legalization
* PvR = Politics vs racial discrimination

## Results

```{r model}
#Model with random slopes for MFQ
Model1_randomslopes <- lmer(MFQTarg ~ MFQSelf_c*Identification_c*Condition + 
                              (MFQSelf_c|Subject), 
                            data = pptnpf1_comb, na.action = na.omit)

tab_model(Model1_randomslopes,
          title = "Effects of predictors and interactions on projection")
```

## Confidence intervals to more than two decimal places
```{r conf int to more decimals}
confint(Model1_randomslopes)
```

# Plots

## Plot of three-way interaction

```{r three-way set-up,include = FALSE, echo = FALSE}
#Creating label for in-group identification so it can be treated like a categorical variable in the plot
describe(pptnpf1_comb$Identification_c)
Ident_label <- c("-1.2" = "Low Identification \n (1 SD Below)",
                 "0" = "Average \n Identification",
                 "1.2" = "High Identification \n (1 SD Above)")

#Getting summary stats to plot
summary.3way <-summary(lsmeans(Model1_randomslopes, 
                               ~MFQSelf_c|Identification_c|Condition, 
                               at = list (MFQSelf_c = c(-3:3), 
                                          Identification_c = c(-1.2, 0, 1.2))))
```

### Model with color

```{r three-way model with color}
#setting figure height and width
knitr::opts_chunk$set(fig.width = 10, fig.height = 10) 

#running the plot
ggplot(summary.3way, aes(x = MFQSelf_c, 
                         y = lsmean, 
                         color = as.factor(Condition))) + 
  facet_grid(.~Identification_c,
             labeller = labeller(Identification_c=Ident_label)) + 
  geom_line(size=1) +
  xlab("MFQ Responses for Self") + 
  ylab("MFQ Responses for Target") + 
  scale_color_discrete("", breaks = c(1,2,3,4), 
                       labels = c ("Politics (P)", 
                                   "Discrimination (D)",
                                   "Marijuana (M)",
                                   "Peanut Butter (PB)")) + 
  ggtitle("Effect of In-group Identification and Issue Polarization \non Projection with Out-Group targets") +
   theme(legend.position = "bottom", 
         legend.text = element_text(size = 11)) +
  theme_minimal(base_size = 13)
```

### Model without color for publication

```{r Three-way BW model for publication}
#running the plot

ggpreview <- function(...) {
  fname <- tempfile(fileext = ".png")
  ggsave(filename = fname, ...)
  system2("open", fname)
  invisible(NULL)
}

knitr::opts_chunk$set(fig.width = 8, fig.height = 6) 

ggplot(summary.3way, aes(x = MFQSelf_c, 
                         y = lsmean, 
                         group = as.factor(Condition))) + 
  facet_grid(.~Identification_c,
             labeller = labeller(Identification_c = Ident_label)) + 
  geom_line(size = .7, aes(linetype = Condition)) +
  xlab("MFQ Responses for Self") + 
  ylab("MFQ Responses for Target") + 
  scale_linetype_manual("Issue polarization", 
                        breaks = c(1,2,3,4), 
                        labels = c ("Political\ncandidate",
                                   "Racial\ndisadvantage",
                                   "Marijuana\nlegalization",
                                   "Peanut\nbutter"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted", 
                                   "dashed")) + 
  ggtitle("Projection predicted by identification and issue polarization") +
  theme_minimal(base_size = 13) +
  theme(legend.text = element_text(size = 10),
        legend.key.width = unit(1, "cm"),
        legend.position = "bottom")
  
ggpreview(width = 8, height = 6, units = "in")
```

## Plot for two-way 

```{r two-way set-up,include = FALSE, echo = FALSE}
#Get summary stats for plot
summary.2way.cond <-summary(lsmeans(Model1_randomslopes, ~MFQSelf_c|Condition, 
                                    at = list(MFQSelf_c = c(-3:3))))
```

### Plot with color

```{r two-way color plot}
knitr::opts_chunk$set(fig.width = 10, fig.height = 10) 
ggplot(summary.2way.cond, aes(x = MFQSelf_c, 
                              y = lsmean, 
                              color = as.factor(Condition)))+
  geom_line(size = 1) +
  xlab("MFQ Responses for Self") + 
  ylab("MFQ Responses for Target") + 
  scale_color_discrete("Issue polarization condition", 
                       breaks = c(1,2,3,4),
                       labels = c("Political candidate",
                                  "Racial disadvantage", 
                                  "Marijuana legalization",
                                  "Peanut Butter preference")) +
  ggtitle("Effect of Issue Polarization on Projection with Out-Group targets") +
  theme(legend.position = "bottom", 
        legend.text = element_text(size = 11)) +
  theme_minimal(base_size = 13)
```

### Plot without color for publication

```{r two-way BW}
knitr::opts_chunk$set(fig.width = 8, fig.height = 6) 

ggplot(summary.2way.cond, aes(x = MFQSelf_c, 
                              y = lsmean, 
                              group = as.factor(Condition)))+
  geom_line(size = 1, aes(linetype = Condition)) +
  xlab("MFQ Responses for Self") + 
  ylab("MFQ Responses for Target") + 
  scale_linetype_manual("Issue polarization", 
                      breaks = c(1,2,3,4),
                      labels = c("Political\ncandidate",
                                  "Racial\ndisadvantage", 
                                  "Marijuana\nlegalization",
                                  "Peanut\nbutter"),
                       values = c("solid",
                                   "twodash", 
                                   "dotted", 
                                   "dashed")) +
  ggtitle("Effect of issue polarization on projection with out-group targets") +
  theme_minimal(base_size = 13) +
  theme(legend.text = element_text(size = 10),
        legend.key.width = unit(1.2, "cm"))


ggpreview(width = 8, height = 6, units = "in")
```

## Plot for main effect of in-group identification

```{r main effect set-up, include = FALSE}
#information to plot main effect of in-group identification
describe(pptnpf1_comb$Identification_c)

effects_ident_main <- effect(term ="MFQSelf_c:Identification_c", 
                             xlevels = list(Identification_c = c(-1.19, 0, 1.19)), 
                             mod = Model1_randomslopes)

effects_ident_main <- as.data.frame(effects_ident_main)

effects_ident_main$Identification <- as.factor(effects_ident_main$Identification_c)
```

### Color plot

```{r}
knitr::opts_chunk$set(fig.width = 8, fig.height = 6) 

ggplot(effects_ident_main, aes(MFQSelf_c, 
                               fit, 
                               group = Identification, 
                               color = Identification)) +
  geom_smooth(method = "lm") +
  scale_color_discrete("Identification", breaks = c(-1.19, 0, 1.19), 
                       labels =c ("Weak Identification \n (1 SD Below)", 
                                  "Average Identification", 
                                  "High Identification \n (1 SD above)")) +
  theme_minimal() +
  labs(title = "Projection predicted by degree of in-group identification",
       x = "MFQ ratings for self",
       y = "MFQ ratings for target")
```

### Plot without color for publication

```{r}
knitr::opts_chunk$set(fig.width = 8, fig.height = 6) 

ggplot(effects_ident_main, aes(MFQSelf_c, 
                               fit, 
                               group = Identification)) +
  geom_line(aes(linetype = Identification), size = .7) +
  scale_linetype_manual("Identification", 
                       breaks = c(-1.19, 0, 1.19), 
                       labels = c ("Low (1 SD below)", 
                                  "Average ", 
                                  "High (1 SD above)"),
                        values = c("solid",
                                   "twodash", 
                                   "dotted")) +
  theme_minimal(base_size = 13) +
  labs(title = "Projection predicted by degree of in-group identification",
       x = "MFQ ratings for self",
       y = "MFQ ratings for target")

ggpreview(width = 8, height = 6, units = "in")
```

